---
title: "Data Preprocessing Notebook"
output: html_notebook
---

```{r}
# Installing required packages
library(quantmod)
library(forecast)
library(tseries)
```

```{r}
# Importing APPLE stock data
getSymbols("AAPL")
stock_data <- Cl(AAPL) # Closing prices
stock_data
```

```{r}
# Column Names
colnames(stock_data)
```

```{r}
# length of the data
length(stock_data)
```

```{r}
# Checking for missing values
missing_indices <- which(is.na(stock_data))
if(length(missing_indices) > 0) {
  print(paste("Number of missing values:", length(missing_indices)))
  # Here, we will use linear interpolation to fill in missing values
  stock_data <- na.approx(stock_data)
} else {
  print("No missing values detected.")
}
```

```{r}
# Checking for NAs values
sum(is.na(diff(log(stock_data))))
```

```{r}
# Removing NAs values
diff_log_data <- diff(log(stock_data))
stock_data_clean <- na.omit(diff_log_data)
stock_data_clean
```

```{r}
# Coerce the time series data to a numeric vector
stock_data_vector <- coredata(stock_data_clean$AAPL.Close)

# Generate a boxplot and determine outliers
boxplot(stock_data_vector, main="Boxplot of Stock Prices")
outliers <- boxplot.stats(stock_data_vector)$out

if(length(outliers) > 0) {
  print(paste("Number of outliers detected:", length(outliers)))
  
  # Get the upper and lower bounds
  upper_bound <- boxplot.stats(stock_data_vector)$stats[5]
  lower_bound <- boxplot.stats(stock_data_vector)$stats[1]
  
  # Cap the outliers
  stock_data_vector[stock_data_vector > upper_bound] <- upper_bound
  stock_data_vector[stock_data_vector < lower_bound] <- lower_bound
  
  # Replace the original column with the capped data
  stock_data_clean$AAPL.Close <- stock_data_vector

} else {
  print("No outliers detected.")
}
```
"Number of outliers detected: 231"

```{r}
# Checking for stationarity
adf.test(stock_data_clean, alternative="stationary", k=0)
```
The result of the Augmented Dickey-Fuller (ADF) test suggests that the diff_log_data_clean series is stationary. Let's break down the results:

Dickey-Fuller Value: The value of -65.454 is the test statistic. The more negative this value is, the stronger the rejection of the hypothesis that there is a unit root.

Lag Order: The lag order is 0, which means no lags of the difference series are included in the test regression. This was set by the k=0 parameter.

p-value: The p-value is given as 0.01. This means that there's strong evidence against the null hypothesis of the series having a unit root (i.e., the series is non-stationary). Since the p-value is very small (and indeed, smaller than the printed p-value as the warning suggests), we can reject the null hypothesis, suggesting that the series is stationary after the differencing.

Warning: The warning "p-value smaller than printed p-value" indicates that the p-value is even smaller than 0.01, which strengthens the evidence against the null hypothesis.

In summary, the test suggests that your log-differenced series is stationary. This is a crucial step when we're preparing a time series for modeling, as many models, like ARIMA, require the input series to be stationary.


```{r}
# Saving the data
saveRDS(diff_log_data_clean, file = "stock_data_clean.rds")
```
















